---
title: "Normal Distribution"
output:
  pdf_document: default
  html_document: default
date: "2024-05-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Normal Probability

Final scores are normally distributed with the mean of 80% and standard deviation of 5%.

X:Final Score

$X\sim N(\mu=80, \sigma =5)$

Calculate the probability that a randomly selected student scored 60% or lower.

$P1=P(X\leq 60)$

```{r}

P1=pnorm(,,)

print(P1)
#library(mosaic)
mosaic::xpnorm(60,80,5)
```
Calculate the probability that a randomly selected student scored between 60% and 80%.

$P2=P(60 \leq X\leq80)$
```{r}
P2= pnorm(,,)-pnorm(,,)
print(P2)
mosaic::xpnorm(c(60,80), 80,5)
```
Calculate:

$P(z>z_{\frac{\alpha}{2}_)=0.05$
```{r}
Z1= qnorm(0.95)
print(Z1)
mosaic::xqnorm(0.95)

```
Calculate the minimum final score of a student who wishes be in the top 1% of the class.

$P(z>z_{\alpha})=0.01$
```{r}
S= qnorm(, , )
print(S)
mosaic::xqnorm(0.99,80,5)
```
85% of a student are scored between 63.59% and x%. Calculate x score.

PP1=$P(z<63.59)$
```{r}
PP1= pnorm(, , )
print(PP1)
Answer=qnorm(PP1+0.85,80,5)
print(Answer)
PP2=1-(PP1+.85)
print(PP2)
PP3=1-PP2
mosaic::xqnorm(c(PP1,PP3),80,5)
```
Consider 10 randomly selected students. What is the chance the their average final scores is below 72%.


$\bar{x}\sim N(\mu=80, \sigma =5)$


$PAV=P(\bar{x}<72)$

```{r}

PAV=pnorm(,,/sqrt())

print(PAV)
#library(mosaic)
mosaic::xpnorm(72,80,5/sqrt(10))
```
Practice the following RCodes:
```{r}
mosaic::xpnorm(c(-1.96,1.96))

mosaic::xpnorm(c(165,175),170,10)

mosaic::xqnorm(c(0.025,0.025))

mosaic::xcnorm(.8)

```
