---
title: "Inferences About the Difference Between Two Population Means"
output:
  pdf_document: default
  html_document: default
date: "2024-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DIETS


Example 1.9.1 Refer to Example 1.16 from the textbook. (Data set: DIETS).

a) Form a 95% confidence interval for the difference between the population mean weight losses for the two diets. Interpret the result.

```{r }
DIETS <-read.csv("DIETS.csv", header=TRUE)
attach(DIETS)
View(DIETS)
head(DIETS)
mean(WTLOSS[DIET==])
summary(DIETS)
#side-by-side box plots
boxplot()
#Labeling 5 number summaries
A=tapply(WTLOSS, DIET,fivenum)
for(i in 1:length(A)){
        text(x=i+0.35,y= A[[i]],labels= A[[i]], pos = 4, offset = 0.7, col = "green")}
```

## a) 95% confidence interval 

```{r}

TEST=t.test()

TEST

df=TEST$parameter
print(df)

#CI (when REGULAR is the first sample)

t.test()
```


## Summary Statistics for each group


```{r}
library(mosaic)
favstats(WTLOSS~DIET)
#or
tapply(WTLOSS,DIET,IQR)
```

## b) Test of Hypothesis

$H_0: \mu_1 -\mu_2 = 0$ or  $\mu_1  =\mu_2$               

$H_a:  \mu_1 -\mu_2 > 0$ or  $\mu_1 > \mu_2$  
```{r}
#Test_Independent_Samples 

test=t.test()
test
```
Graph the rejection region and find the critical value.

```{r}
library(mosaic)

mosaic::xqt(0.95,df)

#Critical Value

print(qt(0.95,df))

#install.packages("gginference")
library(gginference)
ggttest(test)
detach(DIETS)
```



## Pooled t-test

READING

Example 1.9.2 Refer to Example 1.17 from the textbook. (Data set: READING) (Independent samples)

```{r }
READING <-read.csv("READING.csv", header=TRUE)
attach(READING)
View(READING)
head(READING)
summary(READING)

```
a) Use the data in the table to test whether the true mean test scores differ for the new method and the standard method. 
## a) Test of Hypothesis
$H_0: \mu_1 -\mu_2 = 0$ or  $\mu_1  =\mu_2$               

$H_a:  \mu_1 -\mu_2 \ne 0$ or  $\mu_1 \ne \mu_2$


## Checking the Conditions 

```{r }
#side-by-side box plots
boxplot(SCORE~METHOD)
library(mosaic)
favstats(SCORE~METHOD) 

```


```{r}
#Q-Q Plots
par(mfrow=c(1,2))
qqnorm(SCORE[METHOD=="NEW"], main="Normal Q-Q Plot for NEW")
qqline(SCORE[METHOD=="NEW"])
qqnorm(SCORE[METHOD=="STD"], main="Normal Q-Q Plot for STD")
qqline(SCORE[METHOD=="STD"])
```
```{r}
#Test_Independent_Samples_Equal variances_Pooled t-test

test=t.test()
test

```
Graph the rejection region and find the critical value(s).

```{r}
library(mosaic)


mosaic::xct(0.95,20)
#Critical Values

c(-1,1)*qt(.975,20)

library(gginference)
ggttest(test)

detach(READING)
```

## Paired Difference Experiment


Example 1.9.3 Data set: PAIRED shows the data on reading test scores for eight pairs of slow learners with similar reading IQs of which one member is randomly assigned to the standard teaching method while the other is assigned to the new method.



```{r }
PAIRED<- read.csv("PAIRED.csv", header=TRUE)

attach(PAIRED)
head(PAIRED)
```
## a) Test of Hypothesis
$H_0: \mu_d = 0$ or  $\mu_1 -\mu_2 = 0$               

$H_a: \mu_d > 0$ or  $\mu_1 -\mu_2 > 0$  
```{r}
test=t.test(NEW, STANDARD, conf.level=0.95, alternative = c("greater"), paired=TRUE)

```
Graph the rejection region and find the critical value.

```{r}
library(mosaic)


mosaic::xqt(0.95,7)
#Critical Value

print(qt(.95,7))

library(gginference)
ggttest(test)

detach(PAIRED)
```
## Comparing Two Population Variances_F Test

 Example 1.10.1 Refer to Example 1.18 from the textbook. (Data set: READING)
 
 The use of the t statistic was based on the assumption that the population variances of the test scores were equal for the two methods. Conduct a test of hypothesis to check this assumption at $\alpha = .10$.
```{r }
READING <- read.csv("READING.csv")
attach(READING)
head(READING)

```

```{r }
library(mosaic)
favstats() 
var.test()



```
Graph the rejection region and find the critical value.

```{r}
library(mosaic)
mosaic::xqf(0.95,11,9)
#Critical Value

print(qf(.95,11,9))



detach(READING)
```


Example 1.9.1 Refer to Example 1.16 from the textbook. (Data set: DIETS).
 The use of the t statistic was based on the assumption that the population variances of the test scores were not equal for the two diets. Conduct a test of hypothesis to check this assumption at $\alpha = .10$.


```{r }
DIETS <- read.csv("DIETS.csv")
attach(DIETS)
head(DIETS)
var.test()
```

